---
title: "230601_Path_seq_revise_plot"
author: "HuangFei"
date: "2023/6/2"
output: html_document
---

```{r}
set.seed(1000)
library(data.table)
library(Seurat)
library(DOSE)
library(org.Hs.eg.db)
library(clusterProfiler)
library(ggsci)
library(ggpubr)
library(ggplot2)
library(DOSE)
library(AnnotationDbi)
library(pheatmap)
library(RColorBrewer)

my_theme <- theme(plot.title = element_text(hjust = 0.5, size = 17),
                  legend.position = 'right',
                  legend.title =element_text(size=15),
                  legend.text = element_text(size=15),
                  axis.text.x = element_text(size=16),
                  axis.title.x = element_text(size=16),
                  axis.title.y = element_text(size=16),
                  axis.text.y  = element_text(size=16),
                  panel.border = element_blank(),
                  axis.line.x = element_line(size=0.25, color="black"),
                  axis.line.y = element_line(size=0.25, color="black"),
                  panel.grid.minor.x = element_blank(), panel.grid.minor.y = element_blank(),
                  panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
                  panel.background = element_rect(fill='white'),
                  legend.key=element_blank(),
                  strip.text.x = element_text(size=15),
                  strip.text.y = element_text(size=15),
                  strip.background = element_rect(colour = 'white', fill = 'white'))


run_KEGG_human = function(candidate_gene, background_gene = NULL, gene_format = 'SYMBOL', cutoff=0.05,
                    showCategory = 20, font.size = 10,title = 'KEGG enrichment'){
  diff_gene_ID<-clusterProfiler::bitr(candidate_gene, fromType = gene_format, toType ="ENTREZID", OrgDb="org.Hs.eg.db")

  if(is.null(background_gene)){
    ekegg<-enrichKEGG(gene=diff_gene_ID$ENTREZID, organism = 'hsa',
                      pAdjustMethod = "BH",
                      qvalueCutoff =  cutoff,
                      pvalueCutoff = cutoff)
  }else{
    background_gene = clusterProfiler::bitr(background_gene, fromType = gene_format, toType ="ENTREZID", OrgDb="org.Hs.eg.db")
  ekegg<-enrichKEGG(gene=diff_gene_ID$ENTREZID, organism = 'hsa',
                    universe  = background_gene$ENTREZID,
                    pAdjustMethod = "BH",
                    qvalueCutoff =cutoff,
                    pvalueCutoff = cutoff)
  }
  #ekegg <- setReadable(ekegg, OrgDb = org.Mm.eg.db, keytype = "ENTREZID")
  ekegg.table = as.data.frame(ekegg)
  if(nrow(ekegg.table)>0){
    ekegg.table = ekegg.table[order(ekegg.table$p.adjust),]
    print(ekegg.table$Description)
    print(dotplot(ekegg, showCategory = showCategory, font.size = font.size, x='Count', title= title))
  }
  return(ekegg)
}

run_GO_human = function(candidate_gene, background_gene=NULL, gene_format = 'SYMBOL', ontology = 'BP', cutoff=0.05,
                  showCategory=10,font.size=10,title = 'GO enrichment'){
  diff_gene_ID<-clusterProfiler::bitr(candidate_gene, fromType = gene_format, toType ="ENTREZID", OrgDb="org.Hs.eg.db")
  if(is.null(background_gene)){
    ego <-  simplify(enrichGO(gene = diff_gene_ID$ENTREZID,  OrgDb = org.Hs.eg.db,
                    keyType = 'ENTREZID', ont = ontology, readable = T,
                    pAdjustMethod = "BH", qvalueCutoff =   cutoff, pvalueCutoff  = cutoff))
  } else{
    background_gene = clusterProfiler::bitr(background_gene, fromType = gene_format, toType ="ENTREZID", OrgDb="org.Hs.eg.db")
    ego <-  simplify(enrichGO(gene = diff_gene_ID$ENTREZID,  OrgDb = org.Hs.eg.db,
                              universe = background_gene$ENTREZID,
                     keyType = 'ENTREZID', ont = ontology, readable = T,
                     pAdjustMethod = "BH", qvalueCutoff  = cutoff, pvalueCutoff  = cutoff))
  }

  if(nrow(ego@result)>0){
    print(dotplot(ego, showCategory = showCategory, font.size = font.size, x='Count',title=title))
  }
  return(ego)
}
```


```{r}
load("~/Project/ChenLab/result/rdata/Human_patch_part3.RData")
Human_patch = NormalizeData(Human_patch)
head(Human_patch@meta.data)
```


```{r,fig.width=5,fig.height=4.5}
DimPlot(Human_patch, group.by = 'RNA_snn_res.0.6')
```

```{r}
meta.all = read.table('~/Project/ChenLab/result/All patch-seq neurons meta data.txt', header = T, sep = '\t',fill = T)
meta.all#[,96:100]

meta.194 = read.table('~/Project/ChenLab/result/194 patch-seq neurons meta data.txt', header = F, sep = '\t',fill = T)
meta.194

length(intersect(colnames(Human_patch), meta.194$V1))

length(intersect(rownames(Human_patch@meta.data), meta.194$V1))
length(intersect(rownames(Human_patch@meta.data), meta.194$V69))

length(intersect(colnames(Human_patch), meta.all$seq.ID.1))
length(intersect(colnames(Human_patch), meta.all$seq.ID))

meta2use = meta.all[meta.all$seq.ID%in%colnames(Human_patch),]
colnames(meta2use)[18] = 'Layer'
meta2use$age = gsub('m|d','',meta2use$age)
meta2use$age = as.numeric(gsub('y','.',meta2use$age))
summary(meta2use$age)
rownames(meta2use) = meta2use$seq.ID

meta2use = meta2use[colnames(Human_patch),]
```

## DEG identification
```{r}
Human_patch@meta.data = cbind(Human_patch@meta.data, meta2use[,c(7,9,10,13,16:18,20)])

Human_patch@meta.data$SubType = factor(plyr::revalue(Human_patch@meta.data$RNA_snn_res.0.6,
                                      c('1' = 'INT',
                                        '0' = 'PY1',
                                        '2' = 'PY2',
                                        '3' = 'PY3')), levels = c('INT',paste0('PY',1:3)))
```


```{r}
Human_patch@meta.data[grep('other|un',Human_patch@meta.data$Layer),'Layer'] = 'unknown'

Human_patch@meta.data$Layer = factor(Human_patch@meta.data$Layer,
                                     levels = c('upper','deeper','unknown'))

write.csv(tibble::rownames_to_column(Human_patch@meta.data,var = 'cellID'),
          row.names = F, quote = F,
          file = '~/Project/ChenLab/Path-seq/result/text/ChenLab_Pathseq_197_cell_meta_info.csv')
```


```{r}
meta.info = read.csv('~/Project/ChenLab/Path-seq/result/text/V2_ChenLab_Pathseq_197_cell_meta_info.csv', header = T)
meta.info
rownames(meta.info) = meta.info$cellID
meta.info$Donor = substr(meta.info$cellID,1,8)
table(meta.info$Donor)

Human_patch@meta.data = meta.info

table(meta.info$Brain.area)
table(meta.info$pathology)
colnames(meta.info)
write.csv(meta.info[,c(1,27,26,18:25,2:17)],quote = F,row.names = F,
          '~/Project/ChenLab/Path-seq/result/text/V3_ChenLab_Pathseq_197_cell_meta_info.csv')
```


## umap colored by characterization
```{r}
meta.info = meta.info[grep('hipp',meta.info$Brain.area,invert = T),]

brain.area = pal_simpsons()(4); names(brain.area) = sort(unique(meta.info$Brain.area))
pathology = pal_igv()(4); names(pathology) = sort(unique(meta.info$Pathology))
layer = viridis::viridis_pal()(3); names(layer) = sort(unique(meta.info$Layer))

Human_patch = Human_patch[,rownames(meta.info)]
dim(Human_patch)
```


```{r,fig.width=5,fig.height=4.5}
colnames(Human_patch@meta.data)

DimPlot(Human_patch, group.by = 'sex')+
  scale_color_jco()+ labs(title = 'Gender') + my_theme


```

```{r,fig.width=5.2,fig.height=4.5}
Human_patch@meta.data[grep('other|un',Human_patch@meta.data$Layer),'Layer'] = 'unknown'

Human_patch@meta.data$Layer = factor(Human_patch@meta.data$Layer,
                                     levels = c('upper','deeper','unknown'))

DimPlot(Human_patch, group.by = 'Layer')+
  scale_color_manual(values = viridis_pal()(3))+ labs(title = 'Layer') + my_theme

DimPlot(Human_patch, group.by = 'pathology')+
   labs(title = 'Pathology') + my_theme
```

```{r,fig.width=5.5,fig.height=4.5}
DimPlot(Human_patch, group.by = 'Brain.area')+
  scale_color_simpsons()+ labs(title = 'Brain area') + my_theme

```


```{r,fig.width=7,fig.height=4.5}
DimPlot(Human_patch, group.by = 'Donor')+labs(title = 'Donor') +
  scale_color_manual(values = scales::hue_pal()(35)[c(1:3,13,5:12,4,14:35)])+
  my_theme 

```


```{r,fig.width=5,fig.height=4.5}
hmcols <- colorRampPalette(c('#E8C8C4','#B37290','#5C3461'))(50)

tmp = matrix(meta2use$age,nrow = 1)
rownames(tmp) = 'Age'

Human_patch@assays$RNA@data = rbind(Human_patch@assays$RNA@data, tmp)

FeaturePlot(Human_patch,'Age', cols = c('#E8C8C4','#5C3461'), ncol = 1)
```


```{r}
setwd('~/Project/ChenLab/Path-seq/figure/characterization/')

pdf('Gender_colored_UMAP.pdf',5,4.5)
DimPlot(Human_patch, group.by = 'sex')+
  scale_color_jco()+ labs(title = 'Gender') + my_theme
dev.off()

pdf('Age_colored_UMAP.pdf',5,4.5)
FeaturePlot(Human_patch,'Age', cols = c('#E8C8C4','#5C3461'), ncol = 1)
dev.off()


cell2plot = rownames(meta.info[grep('hippo',meta.info$Brain.area,invert = T),])
pdf('Brain_area_UMAP.pdf',5.5,4.5)
DimPlot(Human_patch[,cell2plot], group.by = 'Brain.area')+
  scale_color_simpsons()+ labs(title = 'Brain area') + my_theme
dev.off()

pdf('Pathology_colored_UMAP.pdf',5.2,4.5)
DimPlot(Human_patch, group.by = 'pathology')+
   labs(title = 'Pathology') + my_theme
dev.off()

pdf('Layer_colored_UMAP.pdf',5.2,4.5)
DimPlot(Human_patch, group.by = 'Layer')+
  scale_color_manual(values = viridis_pal()(3))+ labs(title = 'Layer') + my_theme
dev.off()

pdf('Donor_colored_UMAP.pdf',4.5,4.5)
DimPlot(Human_patch, group.by = 'Donor')+labs(title = 'Donor') +
  my_theme + theme(legend.position = 'none')
dev.off()

pdf('~/Project/ChenLab/Path-seq/figure/characterization/V2_Donor_colored_UMAP.pdf',4.5,4.5)
DimPlot(Human_patch, group.by = 'Donor')+labs(title = 'Donor') +
  scale_color_manual(values = scales::hue_pal()(35)[c(1:3,13,5:12,4,14:35)])+
  my_theme + theme(legend.position = 'none')
dev.off()
```



```{r}
Idents(Human_patch) = 'SubType'

cluster.markers = FindAllMarkers(Human_patch, only.pos = F)

setDT(cluster.markers)

cluster.markers[p_val_adj<0.05&pct.1>pct.2]

table(cluster.markers[p_val_adj<0.05&pct.1>pct.2]$cluster)
table(cluster.markers[p_val<0.05&pct.1>pct.2]$cluster)
cluster.markers[,cluster:=factor(cluster,levels = c('INT',paste0('PY',1:3)))]
sig.markers = cluster.markers[p_val_adj<0.05&pct.1>pct.2]
```


### List of interesting genes
```{r}
gene_list = c('GAD1', 'GAD2', 'ERBB4', 'DLX1', 'DLX2', 'LHX6', 'CUX2', 'SLC17A6', 'SLC17A7', 
              'LINC00507', 'SATB2', 'TBR1','TSHZ2', 'DLC1', 'PAK6', 'SLC22A10', 'FOXP1', 'TRHDE',
              'CHST8', 'HS3ST2','RPS6', 'RHEB', 'EIF4E' ,'EGR1', 'EGR2', 'EGR3', 'JUN' ,'DUSP6', 
              'SOX11', 'PTMA', 'RACK1', 'C3', 'CDKN1A', 'NFKBIA' , 'IL1A', 'IL1B', 'IL6' ,'CCL4', 'CCL2',
              'CXCL8', 'CXCL12' , 'MMP2', 'CTSB' , 'ICAM1', 'TNFRSF1B', 'TNFRSF12A', 'PLAUR' ,'IL6', 
              'PTGS2', 'CX3CR1', 'TMEM119', 'CSF1R', 'SELPLG', 'MARCKS', 'IRF8', 'C1QA', 'CD74',
              'HLA-A', 'HLA-B', 'HLA-C', 'HLA-E', 'HLA-DMB',' HLA-DPA1', 'HLA-DQB1', 'UCP2', 'CCND1', 
              'TREM2', 'C1QA', 'APOE', 'CTSC', 'CTSS','NEFH')
```

```{r}
sig.markers[gene%in%gene_list][order(cluster,gene)]

gene2show = c('GAD1', 'GAD2', 'ERBB4', 'DLX1', 'DLX2', 'LHX6',
              'CUX2', 'SLC17A6', 'SLC17A7', 'LINC00507', 'SATB2', 'TBR1','PTGS2', 'RHEB', 'EIF4E' ,
              'RPS6','EGR1', 'EGR2', 'EGR3', 'DUSP6', 'C3', 'CDKN1A',  'CCL2','CD74','C1QA','NFKBIA', 
              'TSHZ2', 'DLC1', 'PAK6', 'SLC22A10', 'FOXP1', 'TRHDE','CHST8', 'HS3ST2','NEFH')
```

```{r,fig.width=5,fig.height=4.5}
color2use = scales::hue_pal()(4)[c(2,1,3,4)]
DimPlot(Human_patch, group.by = 'SubType', cols = color2use)
```


```{r}
path.mat = Human_patch@assays$RNA@data
gene.melt = reshape2::melt(as.matrix(path.mat[rownames(path.mat)%in%gene_list,]))
setDT(gene.melt)
gene.melt
```


## Violin plots
```{r,fig.width=5,fig.height=4}
setwd('~/Project/ChenLab/Path-seq/figure/VlnPlot/')

for(gene in gene_list[grep('HLA',gene_list,invert = T)]){
  ym = max(gene.melt[Var1==gene]$value)
  ym
  vln = VlnPlot(Human_patch,gene,ncol = 1,group.by = 'SubType',
                cols = color2use, slot = 'data', y.max = (ym+0.5)) + 
     stat_compare_means(label.y = (ym+0.2),label.x=2,size=5)  +
    labs(x = '') + theme(legend.position = 'none')
  print(vln)
  pdf(paste0(gene,'_violin_plot.pdf'),5,4)
  print(vln)
  dev.off()
  #assign(gene, vln)
}
```


```{r,fig.width=5,fig.height=5}
gene = 'PTGS2'
ym = max(gene.melt[Var1==gene]$value)
ym

VlnPlot(Human_patch, 'PTGS2', ncol = 1,group.by = 'SubType',
                cols = color2use, slot = 'data',y.max = 3.2)+
  stat_compare_means(comparisons = list(c('INT','PY1'),c('INT','PY2'),c('INT','PY3'),
                                        c('PY1','PY2'),c('PY1','PY3'),c('PY2','PY3')))+
  theme(legend.position = 'none')
```


```{r}
pdf('~/Project/ChenLab/Path-seq/figure/VlnPlot/PTGS2_detailed_pvalue_VlnPlot.pdf',5,5)
VlnPlot(Human_patch, 'PTGS2', ncol = 1,group.by = 'SubType',
                cols = color2use, slot = 'data',y.max = 3.2)+
  stat_compare_means(comparisons = list(c('INT','PY1'),c('INT','PY2'),c('INT','PY3'),
                                        c('PY1','PY2'),c('PY1','PY3'),c('PY2','PY3')))+
  theme(legend.position = 'none')
dev.off()
```

```{r,fig.width=5,fig.height=5}
gene = 'NEFH'
ym = max(gene.melt[Var1==gene]$value)
ym

VlnPlot(Human_patch, 'NEFH', ncol = 1,group.by = 'SubType',
                cols = color2use, slot = 'data',y.max = 2.7)+
  stat_compare_means(label.y = 2.6,label.x=2,size=5)  +
  stat_compare_means(comparisons = list(c('INT','PY1'),c('INT','PY2'),c('INT','PY3'),
                                        c('PY1','PY2'),c('PY1','PY3'),c('PY2','PY3')))+
  theme(legend.position = 'none')
```

```{r}
pdf('~/Project/ChenLab/Path-seq/figure/VlnPlot/NEFH_detailed_pvalue_VlnPlot.pdf',5,5)
VlnPlot(Human_patch, 'NEFH', ncol = 1,group.by = 'SubType',
                cols = color2use, slot = 'data',y.max = 2.7)+
  stat_compare_means(label.y = 2.6,label.x=2,size=5)  +
  stat_compare_means(comparisons = list(c('INT','PY1'),c('INT','PY2'),c('INT','PY3'),
                                        c('PY1','PY2'),c('PY1','PY3'),c('PY2','PY3')))+
  theme(legend.position = 'none')
dev.off()
```

## Feature plots
```{r,fig.width=9,fig.height=7.5}
#FeaturePlot(Human_patch, gene_list[1:9], cols = c('lightgrey','#FF4500'),ncol = 3,order = T)
FeaturePlot(Human_patch, gene_list[1:9], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[10:18], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[19:27], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[28:36], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[37:45], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[46:54], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[55:63], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
FeaturePlot(Human_patch, gene_list[64:71], cols = c('lightgrey','#FF4500'),ncol = 3,order = F)
```


```{r,fig.width=3.8,fig.height=3.5}
setwd('~/Project/ChenLab/Path-seq/figure/FeauturePlot')

for(gene in gene_list[grep('HLA',gene_list,invert = T)]){
  pdf(paste0(gene,'_FeauturePlot.pdf'),3.8,3.5,onefile = T)
  print(FeaturePlot(Human_patch, features = gene, cols = c('lightgrey','#FF4500'),ncol = 1,order = T))
  print(FeaturePlot(Human_patch, features = gene, cols = c('lightgrey','#FF4500'),ncol = 1,order = F))
  dev.off()
}
```

## heatmaps to show DEGs

## Heatmap to show top100 DEGs (p-value < 0.05)
```{r,fig.width=9,fig.height=8}
library(ComplexHeatmap)
#gene2show = gene_list
table(cluster.markers[p_val<0.05&pct.1>pct.2]$cluster)
p05.markers = rbind(cluster.markers[p_val<0.05&pct.1>pct.2][,head(.SD,200), by = .(cluster)],
                    sig.markers[gene%in%c('RHEB', 'EIF4E')])[order(cluster,gene)]
p05.markers[gene%in%gene2show]
#p05.markers = sig.markers[order(cluster,gene)]
p05.markers[,seq:=1:nrow(p05.markers)]
p05.markers[gene%in%gene2show]
meta2use = meta.info
meta2use$Pathology = 'Unknown'
meta2use[grep('FCD',meta2use$pathology),'Pathology'] = 'FCD'
meta2use[grep('TLE',meta2use$pathology),'Pathology'] = 'TLE'
meta2use[grep('PT',meta2use$pathology),'Pathology'] = 'PT'
table(meta2use$Pathology)
meta2use[grep('deep',meta2use$Layer),]$Layer = 'Deep'
meta2use[grep('upp',meta2use$Layer),]$Layer = 'Upper'
table(meta2use$Layer)
dim(meta2use)
mat2use = Human_patch@assays$RNA@data[p05.markers$gene,rownames(meta2use)]
mat2use = t(scale(t(as.matrix(mat2use))))
max(mat2use)
min(mat2use)
dim(mat2use)
mat2use[mat2use>2] = 2
mat2use[mat2use<(-2)] = -2

row_anno = rowAnnotation(foo = anno_mark(at = p05.markers[gene%in%gene2show]$seq,
                                         labels = p05.markers[gene%in%gene2show]$gene))

brain.area = pal_simpsons()(4); names(brain.area) = sort(unique(meta2use$Brain.area))
pathology = pal_igv()(3); names(pathology) = sort(unique(meta2use$Pathology))
layer = viridis::viridis_pal()(3); names(layer) = sort(unique(meta2use$Layer))

col_anno = HeatmapAnnotation(CellType = anno_block(gp = gpar(fill = color2use), labels = c('INT',paste0('PY',1:3))),
                             Brain_area = meta2use$Brain.area, Pathology = meta2use$Pathology, Layer = meta2use$Layer,
                             col = list(Brain_area = brain.area, Pathology = pathology,Layer = layer))

Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        column_split = meta2use$SubType,row_split = p05.markers$cluster,
        col = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
        right_annotation = row_anno, top_annotation = col_anno,show_row_names = F, show_column_names = F)
```

```{r}
pdf('~/Project/ChenLab/Path-seq/figure/characterization/Fig1D_heatmaps_of_CellType_markers.pdf',9,8)
Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        column_split = meta2use$SubType,row_split = p05.markers$cluster,
        right_annotation = row_anno, top_annotation = col_anno,show_row_names = F, show_column_names = F)
dev.off()
```

## GO enrichment analysis
```{r}
sig.markers = cluster.markers[p_val_adj<0.05&pct.1>pct.2]
table(sig.markers$cluster)
p05.markers = cluster.markers[p_val<0.05&pct.1>pct.2]
table(p05.markers$cluster)


PY3.bp = run_GO_human(p05.markers[cluster=='PY3']$gene)

gene_ID<-clusterProfiler::bitr(p05.markers[cluster=='PY3']$gene, fromType = 'SYMBOL',
                               toType ="ENTREZID", OrgDb="org.Hs.eg.db")
gene_ID

write.table(gene_ID$ENTREZID, quote = F, row.names = F, col.names = F,
            file = '~/Project/ChenLab/Path-seq/result/text/PY3_high_DEG_list.txt')
```

```{r}
INT.bp = run_GO_human(p05.markers[cluster=='INT']$gene)

PY1.bp = run_GO_human(p05.markers[cluster=='PY1']$gene)

PY2.bp = run_GO_human(p05.markers[cluster=='PY2']$gene)
```

```{r}
gene_ID<-clusterProfiler::bitr(p05.markers[cluster=='INT']$gene, fromType = 'SYMBOL',
                               toType ="ENTREZID", OrgDb="org.Hs.eg.db")
gene_ID

write.table(gene_ID$ENTREZID, quote = F, row.names = F, col.names = F,
            file = '~/Project/ChenLab/Path-seq/result/text/INT_high_DEG_list.txt')

gene_ID<-clusterProfiler::bitr(p05.markers[cluster=='PY1']$gene, fromType = 'SYMBOL',
                               toType ="ENTREZID", OrgDb="org.Hs.eg.db")


write.table(gene_ID$ENTREZID, quote = F, row.names = F, col.names = F,
            file = '~/Project/ChenLab/Path-seq/result/text/PY1_high_DEG_list.txt')

gene_ID<-clusterProfiler::bitr(p05.markers[cluster=='PY2']$gene, fromType = 'SYMBOL',
                               toType ="ENTREZID", OrgDb="org.Hs.eg.db")


write.table(gene_ID$ENTREZID, quote = F, row.names = F, col.names = F,
            file = '~/Project/ChenLab/Path-seq/result/text/PY2_high_DEG_list.txt')
```


```{r}
INT.bp_df = as.data.table(INT.bp@result)

PY1.bp_df = as.data.table(PY1.bp@result)

PY2.bp_df = as.data.table(simplify(PY2.bp)@result)

INT.bp_df[qvalue<0.05&Count>10]

PY1.bp_df[qvalue<0.05&Count>10]

PY2.bp_df[qvalue<0.01&Count>10]
```

### barplot of GO results
```{r,fig.width=9.2,fig.height=5}
INT.bp_df[,logP:=-log10(pvalue)]

INT_bar = ggbarplot(INT.bp_df[1:10], x = 'Description', y = 'logP', fill = pal_aaas()(2)[2],
        position = position_dodge(), rotate = TRUE,width = 0.7,sort.val =  'asc' ) + 
  geom_hline(yintercept=c(-log10(0.05)),cex=1, colour="red", linetype="dashed")+ylim(c(0,10))+
  labs( x ='',y='-log10(p-value)',title = 'GO analysis of INT DEGs') + my_theme 

INT_bar


PY1.bp_df[,logP:=-log10(pvalue)]

PY1_bar = ggbarplot(PY1.bp_df[1:10], x = 'Description', y = 'logP', fill = pal_aaas()(2)[2],
        position = position_dodge(), rotate = TRUE,width = 0.7,sort.val =  'asc' ) + 
  geom_hline(yintercept=c(-log10(0.05)),cex=1, colour="red", linetype="dashed")+
  labs( x ='',y='-log10(p-value)',title = 'GO analysis of PY1 DEGs') + my_theme 

PY1_bar


PY2.bp_df[,logP:=-log10(pvalue)]

PY2_bar = ggbarplot(PY2.bp_df[c(1,3,5,6,9,12:13,19,25:26)], x = 'Description', y = 'logP', fill = pal_aaas()(2)[2],
        position = position_dodge(), rotate = TRUE,width = 0.7,sort.val =  'asc' ) + 
  geom_hline(yintercept=c(-log10(0.05)),cex=1, colour="red", linetype="dashed")+
  labs( x ='',y='-log10(p-value)',title = 'GO analysis of PY2 DEGs') + my_theme 

PY2_bar
```

```{r,fig.width=9,fig.height=5}
PY3.bp_df = fread('~/Project/ChenLab/Path-seq/result/text/PY3_high_DEG_GO_result.txt',
                  header = T, sep = '\t')
PY3.bp_df[,Description:=mapply(function(x)x[2], strsplit(Term,'~'))]
PY3.bp_df[,logP:=-log10(PValue)]
PY3.bp_df[PValue<0.05&Count>=5,.(Description,PValue,Count)]

PY3_bar = ggbarplot(PY3.bp_df[PValue<0.05&Count>=5][1:10], x = 'Description', y = 'logP', fill = pal_aaas()(2)[2],
        position = position_dodge(), rotate = TRUE,width = 0.7,sort.val =  'asc') + 
  geom_hline(yintercept=c(-log10(0.05)),cex=1, colour="red", linetype="dashed")+ylim(c(0,10))+
  labs( x ='',y='-log10(p-value)',title = 'GO analysis of PY3 cluster DEGs') + my_theme 

PY3_bar
```

```{r}
pdf('~/Project/ChenLab/Path-seq/figure/GO/GO_bar/INT_high_DEG_GO_barplot.pdf',9,5)
INT_bar
dev.off()
pdf('~/Project/ChenLab/Path-seq/figure/GO/GO_bar/PY1_high_DEG_GO_barplot.pdf',9,5)
PY1_bar
dev.off()
pdf('~/Project/ChenLab/Path-seq/figure/GO/GO_bar/PY2_high_DEG_GO_barplot.pdf',9,5)
PY2_bar
dev.off()
pdf('~/Project/ChenLab/Path-seq/figure/GO/GO_bar/PY3_high_DEG_GO_barplot.pdf',9,5)
PY3_bar
dev.off()
```

```{r}
table(p05.markers$cluster)
write.table(p05.markers, sep = ',', row.names = F, quote = F,
            file = '~/Project/ChenLab/Path-seq/result/Path_seq_SubType_DEG_results.csv')
```


## brain region analysis in different subtype
```{r}
meta = Human_patch@meta.data
meta$Pathology = 'FCD'
meta[grep('PT',meta$pathology),'Pathology'] = 'PT'
meta[grep('TLE',meta$pathology),'Pathology'] = 'TLE'
table(meta$Pathology)
meta -> Human_patch@meta.data
```


```{r}
PY1 = subset(Human_patch, SubType=='PY1')
table(PY1$Pathology)
colnames(PY1@meta.data)
Idents(PY1) = 'Pathology'

PY1.DEG = FindAllMarkers(PY1, only.pos = T, logfc.threshold = 0.25)
setDT(PY1.DEG)
table(PY1.DEG$cluster)
table(PY1.DEG[avg_log2FC>log2(1.5)]$cluster)

PY1.DEG[gene%in%PY1.markers$gene]
PY1.DEG[,cluster:=factor(cluster,levels = c("FCD","PT","TLE"))]
PY1.DEG = PY1.DEG[order(cluster)]
table(PY1.DEG[gene%in%PY1.markers$gene]$cluster)
```

```{r,fig.width=6,fig.height=6}
library(ComplexHeatmap)
pathology = pal_igv()(3); names(pathology) = sort(unique(meta$Pathology))
pathology

PY1 = ScaleData(PY1, features = rownames(PY1))

DoHeatmap(PY1, PY1.DEG$gene, group.by = 'Pathology', group.colors = pathology,slot = 'scale.data',
          disp.min = -2, disp.max = 2)

PY1.meta=PY1@meta.data
PY1.meta = PY1.meta[order(PY1.meta$Pathology),]

mat2use = PY1@assays$RNA@data[PY1.DEG$gene,rownames(PY1.meta)]
mat2use = t(scale(t(mat2use)))
max(mat2use)
min(mat2use)
mat2use[1:5,1:5]

mat2use[mat2use>1.5] = 1.5
mat2use[mat2use<(-1.5)] = -1.5

col_anno = HeatmapAnnotation(Pathology = PY1.meta$Pathology,
                             col = list(Pathology = pathology))

PY1.DEG[,group:=cluster]
PY1.DEG[grep('FCD',cluster)]$group = 'FCD (n=25)'
PY1.DEG[grep('PT',cluster)]$group = 'PT (n=114)'
PY1.DEG[grep('TLE',cluster)]$group = 'TLE (n=52)'

Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        row_split = PY1.DEG$group,column_split = PY1.meta$Pathology,
         col = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
        top_annotation = col_anno,show_row_names = F, show_column_names = F)
```


```{r}
pdf('~/Project/ChenLab/Path-seq/revision/figure/PY1_pathology_specific_DEG_heatmaps.pdf',6,6)
Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        row_split = PY1.DEG$group,column_split = PY1.meta$Pathology,
        top_annotation = col_anno,show_row_names = F, show_column_names = F)
dev.off()
```

```{r}
PY2 = subset(Human_patch, SubType=='PY2')
table(PY2$Pathology)
colnames(PY2@meta.data)
Idents(PY2) = 'Pathology'

PY2.DEG = FindAllMarkers(PY2, only.pos = T, logfc.threshold = 0.25)
setDT(PY2.DEG)
table(PY2.DEG$cluster)
table(PY2.DEG[avg_log2FC>log2(1.5)]$cluster)

PY2.DEG[gene%in%PY2.markers$gene]
PY2.DEG[,cluster:=factor(cluster,levels = c("FCD","PT","TLE"))]
PY2.DEG = PY2.DEG[order(cluster)]
table(PY2.DEG[gene%in%PY2.markers$gene]$cluster)
```

```{r,fig.width=6,fig.height=6}
library(ComplexHeatmap)
pathology = pal_igv()(3); names(pathology) = sort(unique(meta$Pathology))
pathology

PY2 = ScaleData(PY2, features = rownames(PY2))

DoHeatmap(PY2, PY2.DEG$gene, group.by = 'Pathology', group.colors = pathology,slot = 'scale.data',
          disp.min = -2, disp.max = 2)

PY2.meta=PY2@meta.data
PY2.meta = PY2.meta[order(PY2.meta$Pathology),]

mat2use = PY2@assays$RNA@data[PY2.DEG$gene,rownames(PY2.meta)]
mat2use = t(scale(t(mat2use)))
max(mat2use)
min(mat2use)
mat2use[1:5,1:5]

mat2use[mat2use>1.5] = 1.5
mat2use[mat2use<(-1.5)] = -1.5

col_anno = HeatmapAnnotation(Pathology = PY2.meta$Pathology,
                             col = list(Pathology = pathology))

PY2.DEG[,group:=cluster]
PY2.DEG[grep('FCD',cluster)]$group = 'FCD (n=3)'
PY2.DEG[grep('PT',cluster)]$group = 'PT (n=160)'
PY2.DEG[grep('TLE',cluster)]$group = 'TLE (n=76)'

Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        row_split = PY2.DEG$group,column_split = PY2.meta$Pathology,
        top_annotation = col_anno,show_row_names = F, show_column_names = F)
```


```{r}
pdf('~/Project/ChenLab/Path-seq/revision/figure/PY2_pathology_specific_DEG_heatmaps.pdf',6,6)
Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        row_split = PY2.DEG$group,column_split = PY2.meta$Pathology,
        top_annotation = col_anno,show_row_names = F, show_column_names = F)
dev.off()
```


```{r,fig.width=6,fig.height=6}
library(ComplexHeatmap)
pathology = pal_igv()(3); names(pathology) = sort(unique(meta$Pathology))
pathology

PY2 = ScaleData(PY2, features = rownames(PY2))

DoHeatmap(PY2, PY2.DEG$gene, group.by = 'Pathology', group.colors = pathology,slot = 'scale.data',
          disp.min = -2, disp.max = 2)

PY2.meta=PY2@meta.data
PY2.meta = PY2.meta[order(PY2.meta$Pathology),]

mat2use = PY2@assays$RNA@data[PY2.DEG$gene,rownames(PY2.meta)]
mat2use = t(scale(t(mat2use)))
max(mat2use)
min(mat2use)
mat2use[1:5,1:5]

mat2use[mat2use>1.5] = 1.5
mat2use[mat2use<(-1.5)] = -1.5

col_anno = HeatmapAnnotation(Pathology = PY2.meta$Pathology,
                             col = list(Pathology = pathology))

PY2.DEG[,group:=cluster]
PY2.DEG[grep('FCD',cluster)]$group = 'FCD (n=3)'
PY2.DEG[grep('PT',cluster)]$group = 'PT (n=160)'
PY2.DEG[grep('TLE',cluster)]$group = 'TLE (n=76)'

Heatmap(mat2use, name = 'Z-scored', cluster_rows = F, cluster_columns = F,
        row_split = PY2.DEG$group,column_split = PY2.meta$Pathology,
        col = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
        top_annotation = col_anno,show_row_names = F, show_column_names = F)
```



```{r}
PT.bp = run_GO_human(PY2.DEG[cluster=='PT']$gene)
TLE.bp = run_GO_human(PY2.DEG[cluster=='TLE']$gene)
```


```{r,fig.width=9,fig.height=7.5}
table(PY2.DEG[gene%in%PY2.markers$gene]$cluster)

PY2.DEG[gene%in%PY2.markers$gene][cluster=='TLE']

VlnPlot(PY2, features = PY2.DEG[gene%in%PY2.markers$gene][cluster=='TLE']$gene,
        group.by = 'Pathology', ncol = 3, pt.size = 0,cols = pal_igv(alpha = 0.9)(3))

enrichplot::dotplot(FCD.bp,title = 'FCD-specific genes in PY1')
```
