---
title: "Integration_of_Chen_and_Allen_Patchseq"
author: "HuangFei"
date: "2023/10/8"
output: html_document
---

```{r}
load("~/project/ChenLab/result/rdata/Human_patch_part3.RData")
Human_patch = NormalizeData(Human_patch)
meta.info = read.csv('~/project/ChenLab/Path-seq/result/text/V2_ChenLab_Pathseq_197_cell_meta_info.csv', header = T)
meta.info
rownames(meta.info) = meta.info$cellID
meta.info$Donor = substr(meta.info$cellID,1,8)
table(meta.info$Donor)

Human_patch@meta.data = meta.info
getwd()
save(Human_patch, file = '~/project/ChenLab/result/rdata/Chen_197_cells_patchseq_seurat_obj.RData')

load('~/project/ChenLab/Path-seq/public_data/Allen_patchseq_data_seurat_obj.RData')
table(Allen.patch$seurat_clusters)
Allen.patch$SubType = Allen.patch$seurat_clusters
```


## Integration of Patch-seq data between Allen and Chen
## Find integration anchors
```{r}
commone.genes = intersect(rownames(Human_patch), rownames(Allen.patch))

Patch.list = list(Human_patch[commone.genes,], Allen.patch[commone.genes,])

names(Patch.list) = c('Chen','Allen')

features <- SelectIntegrationFeatures(object.list = Patch.list)

CM.list <- lapply(Patch.list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})

anchors <- FindIntegrationAnchors(object.list = Patch.list,
                                  reduction = "rpca",dims = 1:30)

## integrate multiple datasets

Patch.combined <- IntegrateData(anchorset = anchors, dims = 1:30)
Patch.combined <- ScaleData(Patch.combined, verbose = F)
```


```{r}
Patch.combined = merge(Human_patch[commone.genes,], Allen.patch[commone.genes,])
Patch.combined@meta.data$Dataset = 'Chen'
Patch.combined@meta.data[colnames(Allen.patch),]$Dataset = 'Allen'

Patch.combined = NormalizeData(Patch.combined)
Patch.combined = FindVariableFeatures(Patch.combined)
Patch.combined = ScaleData(Patch.combined)
```


###PC selections
```{r, fig.height=7.5,fig.width=7.5}
#Patch.combined <- subset(Patch.combined, subset = nFeature_RNA > nGene & percent.mt < mt)
DefaultAssay(Patch.combined) <- 'RNA'
setwd('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/dm')
#setwd('~/Project/Osteoclast/ZYX/Patch.combined/figure/dm')
Patch.combined <- RunPCA(Patch.combined, npcs=50, verbose=F)
Patch.combined = RunHarmony(Patch.combined, group.by.vars = 'Dataset')

pdf('Harmony_heatmaps.pdf',7.5,7.5)
DimHeatmap(Patch.combined, dims = 1:15,  balanced = TRUE,reduction = 'harmony')
DimHeatmap(Patch.combined, dims = 16:30, balanced = TRUE,reduction = 'harmony')
dev.off()
```


```{r,fig.width=5.5,fig.height=5}
setwd('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/dm')
pdf('Harmony_selection.pdf',5.5,5)
ElbowPlot(Patch.combined, ndims = 30,reduction = 'harmony')
  #geom_vline(mapping = aes(xintercept = 11), size = 1, color = 'red', linetype = 'dashed') +
 # labs(title = paste0('PCA of ', ncol(Patch.combined), ' cells'))
dev.off()
```



### Dimension reduction and Clustering
```{r}
# 20 PCs based on elbow plot
npcs <- 1:11
#npcs = 1:21
Patch.combined <- RunUMAP(Patch.combined, dims = npcs, reduction = 'harmony')
#Patch.combined = RunTSNE(Patch.combined, dims = npcs,reduction = 'harmony')
Patch.combined <- FindNeighbors(Patch.combined, dims = npcs, reduction = 'harmony')

Patch.combined <- FindClusters(Patch.combined, resolution = 0.1, reduction = 'harmony')
table(Patch.combined@meta.data$seurat_clusters)
```

### clustering and UMAP
```{r,fig.width=5,fig.height=4.5}
setwd('~/Project/ChenLab/Path-seq/public_data/figure/Patch-seq/dm')
Patch.combined <- FindClusters(Patch.combined, resolution = 0.2)

pdf('Harmony_by_res0.2_cluster.pdf',4.8,4.5)
DimPlot(Patch.combined, label = T, reduction = 'umap',label.size = 5,) + labs(title = '')
dev.off()


pdf('Harmony_by_DataSet.pdf',5,4.5)
DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'Dataset',label.size = 5,) + labs(title = '')
dev.off()

pdf('Harmony_by_SubType.pdf',5,4.5)
set.seed(2020)
DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'SubType',label.size = 5,
        label.box = F, cols = scales::hue_pal()(20)[sample(1:20,9)]) + labs(title = 'Integration of Chen and Allen Patch-seq')
dev.off()
```

```{r,fig.width=5,fig.height=4.5}
DimPlot(Patch.combined, label = T, reduction = 'umap') + labs(title = '')

DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'Dataset') + labs(title = '')

set.seed(2021)
DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'SubType',label.size = 5,
        label.box = F, cols = scales::hue_pal()(20)[sample(1:20,9)]) + labs(title = 'Integration of Chen and Allen Patch-seq')

```

```{r}
save(Patch.combined, file = '~/Project/ChenLab/Path-seq/public_data/Allen_and_Chen_patchseq_integration.RData')
```


## load data
```{r}
load("C:/Users/FeiHuang/Documents/Project/ChenLab/Path-seq/public_data/Allen_an_Chen_patchseq_integration.RData")
Patch.combined@meta.data = Patch.combined@meta.data[,c(1:54,57)]
Patch.combined@meta.data = cbind(Patch.combined@meta.data, as.data.frame(Embeddings(Patch.combined,reduction = 'umap')))
Patch.combined@meta.data
```

## reassign integration clusters
```{r,fig.width=5,fig.height=4.6}
meta = Patch.combined@meta.data
meta[meta$seurat_clusters==0&meta$UMAP_1<4&meta$UMAP_2<0,'seurat_clusters'] = 2
meta$SubType = plyr::revalue(meta$SubType,
                                    c('0'='A1','1'='A2','2'='A3','3'='A4','4'='A5'))
meta[,'Integration_cluster'] = meta$seurat_clusters
meta[meta$seurat_clusters==4,'Integration_cluster'] = 1
meta -> Patch.combined@meta.data

DimPlot(Patch.combined, label = T, reduction = 'umap', group.by = 'Integration_cluster') 
```

### UMAP color by subtype and shape by dataset
```{r,fig.width=5.2,fig.height=4.5}
allen.colors = scales::hue_pal()(10)[c(5,10,7,3,2)]
names(allen.colors) = paste0('A',1:5)
chen.colors = scales::hue_pal()(4)[c(2,1,3,4)]
names(chen.colors) = c('INT',paste0('PY',1:3))

DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'SubType',shape.by =  'Dataset',
        label.size = 5,label.box = F, cols = c(allen.colors,chen.colors)) +labs(title = '')
```


### UMAP split by dataset
```{r,fig.width=9,fig.height=4.5}
DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'SubType',split.by =  'Dataset',
        label.size = 4,label.box = F, cols = c(allen.colors,chen.colors)) +labs(title = '')
```

```{r}
setwd('~/Project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/')
pdf('UMAP_by_Integration_clusters.pdf',4.8,4.6)
DimPlot(Patch.combined, label = T, reduction = 'umap', group.by = 'Integration_cluster') 
dev.off()

pdf('Integration_Allen_and_Chen_Patchseq_UMAP.pdf',5,4.5)
DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'SubType',shape.by =  'Dataset',
        label.size = 5,label.box = F, cols = c(allen.colors,chen.colors)) +labs(title = '')
dev.off()

pdf('Split_Allen_and_Chen_Patchseq_UMAP.pdf',8.5,4.5)
DimPlot(Patch.combined, label = T, reduction = 'umap',group.by = 'SubType',split.by =  'Dataset',
        label.size = 4,label.box = F, cols = c(allen.colors,chen.colors)) +labs(title = '')
dev.off()
```


## focus on PY1-like C0 and PY2-like C2
```{r}
Patch.combined = ScaleData(Patch.combined, features = rownames(Patch.combined))
with(Patch.combined@meta.data, table(seurat_clusters,SubType))

PY2 = subset(Patch.combined, SubType == 'PY2')
PY2@meta.data[order(PY2@meta.data$seurat_clusters),c('seurat_clusters','UMAP_1','UMAP_2')]

C0 = subset(Patch.combined, Integration_cluster == 0)
table(C0$SubType)

C1 = subset(Patch.combined, Integration_cluster == 1)
table(C1$SubType)

C2 = subset(Patch.combined, Integration_cluster == 2)
table(C2$SubType)

C3 = subset(Patch.combined, Integration_cluster == 3)
table(C3$SubType)
```

## cluster fraction/percentage per group
```{r,fig.width=8,fig.height=5}
library(ggpubr)

cell_table = as.data.frame(with(meta, table(SubType, Integration_cluster)))

cell_table
setDT(cell_table)
cell_table = cell_table[Freq>0]
cell_per = cell_table[,100*Freq/sum(Freq), by = .(Integration_cluster)]
cell_per
cell_per$SubType = cell_table$SubType

cell_per

compo_bar1 = ggbarplot(cell_per, x = 'Integration_cluster', y = 'V1', fill = 'SubType',
        position = position_stack(), rotate = TRUE,width = 0.7,palette = c(allen.colors,chen.colors)) + 
  labs( x ='',y='Percentage of cells (%)')  + my_theme + theme(legend.title = element_blank(),
                                                              legend.position = 'right')

compo_bar1


```

```{r,fig.width=8,fig.height=6}
library(ggpubr)

cell_table = as.data.frame(with(meta, table(SubType, Integration_cluster)))


setDT(cell_table)
cell_table = cell_table[Freq>0]
cell_per = cell_table[,100*Freq/sum(Freq), by = .(SubType)]
cell_table[order(SubType)]
cell_per = cell_per[order(SubType)]
cell_per$Integration_cluster = as.character(cell_table[order(SubType)]$Integration_cluster)

cell_per

compo_bar2 = ggbarplot(cell_per, x = 'SubType', y = 'V1', fill = 'Integration_cluster',
        position = position_stack(), rotate = TRUE,width = 0.7) + 
  labs( x ='',y='Percentage of cells (%)')  + my_theme + theme(legend.title = element_blank(),
                                                              legend.position = 'right')

compo_bar2
```

## GSVA estimation of Chen Lab PY1/2 markers
```{r}
chen.markers = fread('~/project/ChenLab/Path-seq/result/Path_seq_SubType_DEG_results.csv', header = T, sep = ',')
table(chen.markers[pct.1>pct.2&p_val_adj<0.05&avg_log2FC>log2(1.5)]$cluster)

sig.markers = chen.markers[pct.1>pct.2&p_val<0.05&avg_log2FC>log2(2)]
#sig.markers = sig.markers[,head(.SD,100),by=.(cluster)]
table(sig.markers$cluster)
sig.markers[avg_log2FC>1&cluster=='PY2']

gene.list = list(GSVA.INT = sig.markers[cluster=='INT']$gene,
                 GSVA.PY1 = sig.markers[cluster=='PY1']$gene,
                 GSVA.PY2 = sig.markers[cluster=='PY2']$gene,
                 GSVA.PY3 = chen.markers[pct.1>pct.2&p_val<0.05&avg_log2FC>log2(1.5)][cluster=='PY3']$gene[1:100])
```

```{r}
library(scales)
library(GSVA)
set.seed(123)

gene.list = list(GSVA.INT = sig.markers[cluster=='INT']$gene,
                 GSVA.PY1 = sig.markers[cluster=='PY1'&p_val_adj<0.05]$gene,
                 GSVA.PY2 = sig.markers[cluster=='PY2']$gene,
                 GSVA.PY3 = chen.markers[pct.1>pct.2&p_val<0.05&avg_log2FC>log2(1.5)][cluster=='PY3']$gene[1:100])


mat = Patch.combined@assays$RNA@data
  
  chen.GSVA <- gsva(as.matrix(mat[rownames(mat)%in%unlist(gene.list),]), gset.idx.list = gene.list, kcdf="Gaussian",method = "gsva", parallel.sz=1)
  #chen.GSVA <- gsva(mat, gset.idx.list = gene.list, kcdf="Poisson",method = "ssgsea", parallel.sz=1)
  chen.GSVA.mat = t(apply(chen.GSVA,1,rescale))
  Patch.combined@assays$RNA@data = rbind(Patch.combined@assays$RNA@data,chen.GSVA.mat)
```

```{r,fig.width=14,fig.height=3}
int = VlnPlot(Patch.combined,c('GSVA.INT'),group.by = 'Integration_cluster',  ncol = 1, pt.size = 0)+ labs(x='',y='GSVA scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

py1 = VlnPlot(Patch.combined,c('GSVA.PY1'),group.by = 'Integration_cluster', ncol = 1, pt.size = 0)+  labs(x='',y='GSVA scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

py2 = VlnPlot(Patch.combined,c('GSVA.PY2'),group.by = 'Integration_cluster', ncol = 1, pt.size = 0)+  labs(x='',y='GSVA scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

py3 = VlnPlot(Patch.combined,c('GSVA.PY3'),group.by = 'Integration_cluster', ncol = 1, pt.size = 0)+  labs(x='',y='GSVA scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

FeaturePlot(Patch.combined, paste0('GSVA.',c('INT',paste0('PY',1:3))),
            reduction = 'umap',cols = colorRampPalette(c('grey','lightgrey','#FF4500'))(10),ncol = 4, order = T,slot = 'data')

FeaturePlot(Patch.combined, paste0('GSVA.',c('INT',paste0('PY',1:3))),
            reduction = 'umap',cols = colorRampPalette(c('grey','lightgrey','#FF4500'))(10),ncol = 4, order = F,slot = 'data')

```


```{r}
pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/Chen_GSVA_score_VlnPlot_across_Integration_cluster.pdf',16,3)
plot_grid(int,py1,py2,py3, ncol = 4)
dev.off()

pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/Chen_GSVA_score_mapping_to_Integration_cluster.pdf',14,3)
FeaturePlot(Patch.combined, paste0('GSVA.',c('INT',paste0('PY',1:3))),
            reduction = 'umap',cols = viridis::viridis(10),ncol = 4, order = F,slot = 'data')
dev.off()


pdf('~/project/ChenLab/Path-seq/figure/AI/Fig2C_Chen_GSVA_score_mapping_to_Integration_cluster.pdf',14,3)
FeaturePlot(Patch.combined, paste0('GSVA.',c('INT',paste0('PY',1:3))),
            reduction = 'umap',cols = colorRampPalette(c('grey','lightgrey','#FF4500'))(10),ncol = 4, order = F,slot = 'data')
dev.off()
```

## AUCell score
```{r,fig.width=5,fig.height=4.5}
library(AUCell)
thrP = 0.01
PopPer = 0.1
dim(mat)
cells_rankings <- AUCell_buildRankings(mat)

names(gene.list) = gsub('GSVA','AUC',names(gene.list))

set.seed(123)

setwd('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration')
pdf('Integration_AUCell_distribution.pdf',5,4.5)
cells_AUC <- AUCell_calcAUC(gene.list, cells_rankings, aucMaxRank=nrow(cells_rankings)*0.05)
cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, nCores=1, 
                                             assign=TRUE,thrP=thrP,smallestPopPercent = PopPer)
dev.off()

selectedThresholds <- getThresholdSelected(cells_assignment)
```


### We assigned celltypes to the cells holding the AUC scors over thershold !
```{r}
### obtain UMAP projection to plot
cellsReduction = as.data.frame(Embeddings(object = Patch.combined, reduction = "umap"))

setwd('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration')
pdf('Integration_AUCell_FeaturePlot.pdf',5,4)
for(geneSetName in names(selectedThresholds)){
  
  # Split cells according to their AUC value for the gene set
  passThreshold <- getAUC(cells_AUC)[geneSetName,] >  selectedThresholds[geneSetName]

    aucSplit <- split(getAUC(cells_AUC)[geneSetName,], passThreshold)
    
    cellsReduction$AUC = getAUC(cells_AUC)[geneSetName,]
    
    
    # Plot
    #plot(cellsReduction[order(cellsReduction$AUC),1:2], main=geneSetName,
    #sub="Pink/red cells pass the threshold",
    #col=cellColor[rownames(cellsReduction[order(cellsReduction$AUC),])], pch=16) 
    
    toplot = cellsReduction[order(cellsReduction$AUC),]
   # if(geneSetName == 'PY2'){
  #        toplot[toplot$AUC>0.4,]$AUC = 0.4
   # }

    umap_cell = ggplot(toplot,
                        aes(x=UMAP_1,y=UMAP_2,col=AUC))+ geom_point(size=0.5)+
       #scale_color_gradientn(colors = colorRampPalette(c('lightgrey', '#FF4500'))(20),limits = c(0,0.5), breaks = seq(0,0.5,length=6)) +
      scale_color_gradient2(low = viridis::viridis(3)[1],
                            mid = viridis::viridis(3)[2],
                            high = viridis::viridis(3)[3]) +
      #scale_color_gradient(low = 'lightgrey', high = '#FF4500') + 
      my_theme+ggtitle(paste0(geneSetName,'\ncutoff: > ',signif(selectedThresholds[geneSetName],2)))
     
    # print(umap_cell)
     assign(geneSetName,umap_cell)
     if(sum(passThreshold) >0 ){
       
    passed = names(passThreshold[passThreshold>0])
    cellsReduction$Identification = 'No'
    cellsReduction[passed,]$Identification = 'YES'
    
    table(cellsReduction$Identification)
    
    umap_ident =  ggplot(cellsReduction[order(cellsReduction$AUC),], aes(x=UMAP_1,y=UMAP_2,col=Identification))+
      geom_point(size=0.5)+scale_color_manual(values = c('lightgrey','#E64B35FF')) + my_theme+
      ggtitle(paste0(geneSetName,'\ncutoff: > ',signif(selectedThresholds[geneSetName],2)))
     
    
    print(umap_ident)
  }
} 
dev.off()

AUC.mat = t(apply(getAUC(cells_AUC),1,rescale))
dim(AUC.mat)
Patch.combined@assays$RNA@data = rbind(Patch.combined@assays$RNA@data,AUC.mat)
```

```{r,fig.width=14,fig.height=3}
int = VlnPlot(Patch.combined,c('AUC.INT'),group.by = 'Integration_cluster',  ncol = 1, pt.size = 0)+ labs(x='',y='AUC scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

py1 = VlnPlot(Patch.combined,c('AUC.PY1'),group.by = 'Integration_cluster', ncol = 1, pt.size = 0)+  labs(x='',y='AUC scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

py2 = VlnPlot(Patch.combined,c('AUC.PY2'),group.by = 'Integration_cluster', ncol = 1, pt.size = 0)+  labs(x='',y='AUC scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

py3 = VlnPlot(Patch.combined,c('AUC.PY3'),group.by = 'Integration_cluster', ncol = 1, pt.size = 0)+  labs(x='',y='AUC scores')+geom_boxplot(width=0.15,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/Chen_AUC_score_VlnPlot_across_Integration_cluster.pdf',16,3)
plot_grid(int,py1,py2,py3, ncol = 4)
dev.off()

pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/Chen_AUC_score_mapping_to_Integration_cluster.pdf',14,3)
FeaturePlot(Patch.combined, paste0('AUC.',c('INT',paste0('PY',1:3))),
            reduction = 'umap',cols = viridis::viridis(10),ncol = 4, order = F,slot = 'data')
dev.off()


```


```{r,fig.width=15,fig.height=6}
VlnPlot(Patch.combined, c('ACTB','GAPDH','CUX2','PTGS2','RPS6','CDKN1A'),
        group.by =  'SubType',ncol = 3, slot = 'data',pt.size = 0,
        cols = c(allen.colors,chen.colors))

VlnPlot(Patch.combined, c('ACTB','GAPDH','CUX2','RPS6'),
        group.by =  'SubType',ncol = 2, slot = 'scale.data',pt.size = 0,
        cols = c(allen.colors,chen.colors))
```


# to show representative genes 
## PY2 markers
```{r,fig.width=12,fig.height=7.5}
VlnPlot(C2, c('C3','C1QA','CCL2','CDKN1A','NFKBIA','RPS6','EIF4E','EGR2','RHEB'),
        group.by =  'SubType',ncol = 3, slot = 'data',pt.size = 0,
        cols = c(allen.colors[paste0('A',c(1,2,4))],chen.colors[paste0('PY',1:2)]))
```


```{r,fig.width=4,fig.height=3}
C2 = subset(C2, SubType!='A4')
VlnPlot(Patch.combined, 'RPS6', group.by =  'Integration_cluster',ncol = 1, slot = 'data',pt.size = 0)+
  geom_boxplot(width=0.1,fill='grey',outlier.shape = NA) + labs(x='') + 
   #stat_compare_means(comparisons = list(c('0','2'),c('1','2'),c('3','2')),font.label=list(size=14))+
  theme(axis.text.x = element_text(size=16,angle = 45,hjust = 0.99,vjust = 0.99),legend.position = 'none')

VlnPlot(C2, 'RPS6', group.by =  'SubType',ncol = 1, slot = 'data',pt.size = 0,
        cols = c(allen.colors[paste0('A',c(1,2,4))],chen.colors[paste0('PY',1:2)]))+
  geom_boxplot(width=0.1,fill='grey',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 45,hjust = 0.99,vjust = 0.99),legend.position = 'none')
```



```{r}
spikie.mat =  Matrix::Matrix(data = as.matrix(ISnorm_res$normalized), sparse = TRUE)
class(spikie.mat)

Patch.combined@assays$RNA@data = spikie.mat
```


```{r,fig.width=3.8,fig.height=3}
for(ge in c('C3','C1QA','CCL2','CDKN1A','NFKBIA','RPS6','EIF4E','EGR2','EGR3','JUN','RHEB','TPI1','ENO1','PGK1','PTGS2','TP53','IL1A','IL1B','CCL4','CCL2','CXCL8','CXCL12','LMNB1')){
# v1= VlnPlot(Patch.combined, ge, group.by =  'Integration_cluster',ncol = 1, slot = 'data',pt.size = 0)+
#  geom_boxplot(width=0.1,fill='white',outlier.shape = NA) + labs(x='') + 
   #stat_compare_means(comparisons = list(c('0','2'),c('1','2'),c('3','2')),font.label=list(size=14))+
#  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')
 v2 = VlnPlot(C2, ge, group.by =  'SubType',ncol = 1, slot = 'data',pt.size = 0,
        cols = c(allen.colors[paste0('A',c(1,2))],chen.colors[paste0('PY',1:2)]))+
  geom_boxplot(width=0.1,fill='grey',outlier.shape = NA) + labs(x='') + 
   my_theme+
  theme(axis.text.x = element_text(size=16,angle = 45,hjust = 0.99,vjust = 0.99),legend.position = 'none')
 #assign(paste0(ge,'.all'),v1)
 print(v2)
 assign(paste0(ge,'.C2'),v2)
}
```


```{r,fig.width=10,fig.height=7.5}
pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/PY2_marker_genes_across_Integration_clusters.pdf',10,7.5)
plot_grid(RPS6.all,C3.all,C1QA.all,CCL2.all,CDKN1A.all,NFKBIA.all,EIF4E.all,EGR2.all,RHEB.all,ncol = 3)
dev.off()
```

```{r,fig.width=10,fig.height=7.5}
pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/PY2_marker_genes_in_cluster2.pdf',10,7.5)
plot_grid(RPS6.C2,C3.C2,C1QA.C2,CCL2.C2,CDKN1A.C2,NFKBIA.C2,EIF4E.C2,EGR2.C2,RHEB.C2,ncol = 3)
dev.off()
```

## PY1 markers
```{r}
for(ge in c('CUX2','PTGS2','SATB2','SLC17A6','SLC17A7','LGI4')){
 v1= VlnPlot(Patch.combined, ge, group.by =  'Integration_cluster',ncol = 1, slot = 'data',pt.size = 0)+
  geom_boxplot(width=0.1,fill='white',outlier.shape = NA) + labs(x='') + 
   #stat_compare_means(comparisons = list(c('0','2'),c('1','2'),c('3','2')),font.label=list(size=14))+
  theme(axis.text.x = element_text(size=16,angle = 0,hjust = 0.99,vjust = 0.99),legend.position = 'none')

 v2 = VlnPlot(C0, ge, group.by =  'SubType',ncol = 1, slot = 'data',pt.size = 0,
        cols = c(allen.colors[paste0('A',c(1:5))],chen.colors[paste0('PY',1:3)]))+
  geom_boxplot(width=0.1,fill='white',outlier.shape = NA) + labs(x='') + 
  theme(axis.text.x = element_text(size=16,angle = 45,hjust = 0.99,vjust = 0.99),legend.position = 'none')
 assign(paste0(ge,'.all'),v1)
 assign(paste0(ge,'.C0'),v2)
}
```


```{r,fig.width=10,fig.height=5}
pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/PY1_marker_genes_across_Integration_clusters.pdf',10,5)
plot_grid(CUX2.all,PTGS2.all,SATB2.all,SLC17A6.all,SLC17A7.all,LGI4.all,ncol = 3)
dev.off()
```

```{r,fig.width=12,fig.height=5}
pdf('~/project/ChenLab/Path-seq/public_data/figure/Patch-seq/Integration/PY1_marker_genes_in_cluster2.pdf',12,5)
plot_grid(CUX2.C0,PTGS2.C0,SATB2.C0,SLC17A6.C0,SLC17A7.C0,LGI4.C0,ncol = 3)
dev.off()
```

## try to compare PY2 genes group by Subtypes, split by integration cluster
```{r}
VlnPlot(Patch.combined, c('RPS6'), split.by =  'SubType',ncol = 1,group.by = 'Integration_cluster', slot = 'data',pt.size = 0,split.plot = T,
        cols = c(allen.colors[paste0('A',c(1,2,4))],chen.colors[paste0('PY',1:2)])) + 
   geom_boxplot(width=0.1,fill='white',outlier.shape = NA) 
```
Split violin is only supported for <3 groups, using multi-violin.




## normalization using ERCC
```{r}
library(ISnorm)
library(dbscan)

dim(mat)
gene_dis<-calculate.dis(as.matrix(mat),detection_rate=0.90)
spike_candidate<-dbscan.pick(dis=gene_dis,ngene=(1:floor(nrow(gene_dis)/25))*5)

spike.res <- candidate.norm(mat=as.matrix(mat),spike_candidate=spike_candidate)

spike.res$spike[[9]]

ISnorm_res<-opt.candidate(mat=as.matrix(mat),candidate_res=spike.res)

dim(ISnorm_res$normalized)

mat[1:6,1:6]
ISnorm_res$normalized[1:6,1:6]
```

```{r}
Patch.mat = ISnorm_res$normalized
Patch.melt = melt(as.matrix(Patch.mat[,]))

head(Patch.melt)
colnames(Patch.melt)[1:2] = c('gene','cellID')

Patch.combined@meta.data$cellID = rownames(Patch.combined@meta.data)
colnames(Patch.combined@meta.data)

length(intersect(Patch.melt$cellID, Patch.combined@meta.data$cellID))

Patch.melt = merge(Patch.melt, Patch.combined@meta.data[,c('cellID','Integration_cluster','SubType','Dataset')],
                   by = 'cellID')

setDT(Patch.melt)
Patch.melt[,SubType:=factor(SubType,levels = c(names(allen.colors),names(chen.colors)))]
```

```{r,fig.width=5,fig.height=3}
for(i in c('RPS6','ACTB','GAPDH')){
   vp = ggviolin(Patch.melt[gene == i], x = 'SubType', y = 'value',fill = 'SubType',width = 1.5,
                palette = c(allen.colors,chen.colors),trim = T) +
    geom_boxplot(width=0.1,fill='white',outlier.shape = NA) + 
  labs(x = '', y = 'Relative expression', title = i) +#  coord_cartesian(ylim=c(0,10)) 
   my_theme + theme(legend.position = 'none')
  print(vp)
}
```


```{r,fig.width=12,fig.height=3}
for(i in c('RPS6','ACTB','GAPDH')){
   vp = ggviolin(Patch.melt[gene == i][SubType%in%c(paste0('A',c(1,2,4)),paste0('PY',1:2))], x = 'SubType', y = 'value',fill = 'SubType',width = 1.2,
                palette = c(allen.colors[paste0('A',c(1,2,4))],chen.colors[paste0('PY',1:2)]),trim = T) +
    geom_boxplot(width=0.1,fill='white',outlier.shape = NA) + 
   facet_wrap(~Integration_cluster,ncol = 4)+
  labs(x = '', y = 'Relative expression', title = i) +#  coord_cartesian(ylim=c(0,10)) 
   my_theme + theme(legend.position = 'none')
  print(vp)
}

```
